---
title: "HW02"
author: "Tyler Allan"
date: "2/5/2021"
output: 
  html_document: default
    theme: paper
  github_document: default
---

<!-- BMW: in the YAML metadata for a document, suboptions for a parameter (like keep_md: true, which is an option for html_document) need to be indented under the parameter. See like what I did with theme: up there now. --> 

```{r setup, include = FALSE}
# BMW: Name the chunk where you load packages 'setup'
knitr::opts_chunk$set(echo = TRUE)
library(gapminder)
library(tidyverse) # BMW: library(tidyverse) already loads dplyr and ggplot2, so you don't need both.
library(ggplot2)
library(dplyr)
library(palmerpenguins) # BMW: Put all library() calls at the top of your script.
```

# 1.1 Use filter() to subset the gapminder data to three countries of your choice in the 1970's

<!-- BMW: In Markdown, you need a space after # in headings -->

<!-- BMW: In Markdown, you need a blank line between paragraphs. A single line break is ignored. -->

```{r}
acjminder <- gapminder %>% 
  filter(country %in% c("Afghanistan","Canada","Japan") & year > 1969 & year < 1980)
print(acjminder)
```


# 1.2 Use the pipe operator to select "country" and "gdpPercap" from your filtered dataset in 1.1

``` {r}
acjminder %>% 
  select(country, gdpPercap)
```

# 1.3 Make a new variable in gapminder for the change in life expectancy from the previous measurement. Filter this table to show all of the entries that have experienced a drop in life expectancy. Hint: you might find the lag() and diff() functions useful.

```{r}
deltalifeExp <- gapminder %>% 
  group_by(country) %>% # BMW: You need to group by country so that you don't compare 2007 for one country to 1952 for the next
  mutate(lagLifeExp = lag(lifeExp, order_by = year))
# BMW: Your arrange() call here didn't do anything because you didn't assign the result to anything.
deltalifeExp %>% 
  filter(lagLifeExp > lifeExp) # BMW: You had this backwards. This would have shown all rows where the life expectancy increased, not droped.
```

# 1.4 Filter gapminder so that it shows the max GDP per capita experienced by each country. Hint: you might find the max() function useful here.

```{r}
gapminder %>% 
  group_by(country) %>% 
  filter(gdpPercap == max(gdpPercap)) # BMW: Here, you can use == because you are comparing against a single value (though the result is the same as with %in%)
```

# 1.5 Produce a scatterplot of Canada's life expectancy vs. GDP per capita using ggplot2, without defining a new variable. That is, after filtering the gapminder data set, pipe it directly into the ggplot function. In your plot, put GDP per capita on a log scale. 

```{r}
gapminder %>% 
  filter(country == "Canada") %>% 
  ggplot() +
  aes(x = lifeExp,
      y = gdpPercap) +
  geom_point() + 
  scale_y_log10() # BMW: For readability, put all plot commands on a new line
```

# 2.1 Pick two quantitative variables to explore. Make a table of descriptive statistics for these variables using summarize (include whatever statistics you feel appropriate). Make a scatterplot of these variables using ggplot()

```{r}
# palmerpenguins::penguins # BMW: Don't print out whole datasets like this in your final scripts.

penguin <- na.omit(penguins)
# penguin # BMW: Don't print out whole datasets like this in your final scripts.
penguin %>%
  summarize(across(c(bill_length_mm, bill_depth_mm),
                   list(mean = mean, SD = sd)))

ggplot(penguins) +
  aes(x = bill_length_mm,
      y = bill_depth_mm) +
  geom_point()
```

# 2.2 Pick one categorical variable and one quantitative variable to explore. Make a summary table giving the sample size (hint: n()) and descriptive statistics for the quantitative variable by group. Make one or more useful plots to visualize these variables.

```{r}
penguin %>%
  group_by(species) %>%
  summarize(
    meangd = mean(bill_depth_mm),
    sdgd = sd(bill_depth_mm),
    meangl = mean(bill_length_mm),
    sdgl = sd(bill_length_mm),
    meangf = mean(flipper_length_mm),
    sdgf = sd(flipper_length_mm),
    n = n()
  )

ggplot(penguin) +
  aes(x = bill_length_mm,
      y = bill_depth_mm,
      fill = species,
      color = species) +
  geom_point()

ggplot(penguin) +
  aes(x = species) +
  geom_bar()

ggplot(penguin) +
  aes(x = bill_length_mm,
      y = flipper_length_mm,
      fill = species,
      color = species) +
  geom_point()

```
