---
title: "HW02"
author: "Tyler Allan"
date: "2/5/2021"
output: html_document: default
        github_document: default
---
```{r}
library(gapminder)
library(tidyverse)
library(ggplot2)
library(dplyr)
```

#1.1 Use filter() to subset the gapminder data to three countries of your choice in the 1970's
```{r}
acjminder <- gapminder %>% 
  filter(country %in% c("Afghanistan","Canada","Japan") & year > 1969 & year < 1980)
print(acjminder)
```


#1.2 Use the pipe operator to select "country" and "gdpPercap" from your filtered dataset in 1.1
``` {r}
acjminder %>% 
  select(country, gdpPercap)
```

#1.3 Make a new variable in gapminder for the change in life expectancy from the previous measurement. Filter this table to show all of the entries that have experienced a drop in life expectancy. Hint: you might find the lag() and diff() functions useful.
```{r}
deltalifeExp <- gapminder %>% 
  mutate(deltalifeExp = lag(lifeExp, order_by = year))
arrange(deltalifeExp, year)
deltalifeExp %>% 
  filter(deltalifeExp < lifeExp)
```

#1.4 Filter gapminder so that it shows the max GDP per capita experienced by each country. Hint: you might find the max() function useful here.
```{r}
gapminder %>% 
  filter(max(gdpPercap))
```

#1.5 Produce a scatterplot of Canada's life expectancy vs. GDP per capita using ggplot2, without defining a new variable. That is, after filtering the gapminder data set, pipe it directly into the ggplot function. In your plot, put GDP per capita on a log scale. 
```{r}

```

#2.1 Pick two quantitative variables to explore. Make a table of descriptive statistics for these variables using summarize (include whatever statistics you feel appropriate). Make a scatterplot of these variables using ggplot()
```{r}

```

#2.2 Pick one categorical variable and one quantitative variable to explore. Make a summary table giving the sample size (hint: n()) and descriptive statistics for the quantitative variable by group. Make one or more useful plots to visualize these variables.
```{r}

```























































