---
title: "HW02"
author: "Tyler Allan"
date: "2/5/2021"
output: 
    github_document: default
    html_document: default
    keep_md: TRUE
---
```{r}
library(gapminder)
library(tidyverse)
library(ggplot2)
library(dplyr)
```

#1.1 Use filter() to subset the gapminder data to three countries of your choice in the 1970's
```{r}
acjminder <- gapminder %>% 
  filter(country %in% c("Afghanistan","Canada","Japan") & year > 1969 & year < 1980)
print(acjminder)
```


#1.2 Use the pipe operator to select "country" and "gdpPercap" from your filtered dataset in 1.1
``` {r}
acjminder %>% 
  select(country, gdpPercap)
```

#1.3 Make a new variable in gapminder for the change in life expectancy from the previous measurement. Filter this table to show all of the entries that have experienced a drop in life expectancy. Hint: you might find the lag() and diff() functions useful.
```{r}
deltalifeExp <- gapminder %>% 
  mutate(deltalifeExp = lag(lifeExp, order_by = year))
arrange(deltalifeExp, year)
deltalifeExp %>% 
  filter(deltalifeExp < lifeExp)
```

#1.4 Filter gapminder so that it shows the max GDP per capita experienced by each country. Hint: you might find the max() function useful here.
```{r}
gapminder %>% 
  group_by(country) %>% 
  filter(gdpPercap %in% max(gdpPercap))
```

#1.5 Produce a scatterplot of Canada's life expectancy vs. GDP per capita using ggplot2, without defining a new variable. That is, after filtering the gapminder data set, pipe it directly into the ggplot function. In your plot, put GDP per capita on a log scale. 
```{r}
gapminder %>% 
  filter(country == "Canada") %>% 
  ggplot() +
  aes(x = lifeExp,
      y = gdpPercap) +
  geom_point() + scale_y_log10()
```

#2.1 Pick two quantitative variables to explore. Make a table of descriptive statistics for these variables using summarize (include whatever statistics you feel appropriate). Make a scatterplot of these variables using ggplot()
```{r}
library(palmerpenguins)
palmerpenguins::penguins

penguin <- na.omit(penguins)
penguin
penguin %>% 
  summarize(meanl = mean(bill_length_mm),
            sdl = sd(bill_depth_mm),
            meand = mean(bill_length_mm),
            sdd = sd(bill_depth_mm),
  )

    ggplot(penguins) +
      aes(x = bill_length_mm,
          y = bill_depth_mm) +
      geom_point()
```

#2.2 Pick one categorical variable and one quantitative variable to explore. Make a summary table giving the sample size (hint: n()) and descriptive statistics for the quantitative variable by group. Make one or more useful plots to visualize these variables.
```{r}
penguin %>% 
  group_by(species) %>% 
  summarize(meangd = mean(bill_depth_mm),
            sdgd = sd(bill_depth_mm),
            meangl = mean(bill_length_mm),
            sdgl = sd(bill_length_mm),
            meangf = mean(flipper_length_mm),
            sdgf = sd(flipper_length_mm),
            n = n())
  ggplot(penguin) +
      aes(x = bill_length_mm,
          y = bill_depth_mm,
          fill = species,
          color = species) +
      geom_point()
  ggplot(penguin) +
    aes(x = species) +
   geom_bar()
  ggplot(penguin) +
  aes(x = bill_length_mm,
      y = flipper_length_mm,
      fill = species,
      color = species) +
  geom_point()

```























































