---
title: "HW02"
author: "Tyler Allan"
date: "2/5/2021"
output: html_document: default
        github_document: default
---
```{r}
library(gapminder)
library(tidyverse)
library(ggplot2)
library(dplyr)
```

#1.1 Use filter() to subset the gapminder data to three countries of your choice in the 1970's
```{r}
transform(gapminder, year = as.factor(year))


acjminder <- gapminder %>% 
  filter(country == c("Afghanistan","Canada","Japan"))
print(acjminder)
```

#1.2 Use the pip operator to select "country" and "gdpPercap" from your filtered dataset in 1.1

#1.3 Make a new variable in gapminder for the change in life expectancy from the previous measurement. Filter this table to show all of the entries that have experienced a drop in life expectancy. Hint: you might find the lag() and diff() functions useful.

#1.4 Filter gapminder so that it shows the max GDP per capita experienced by each country. Hint: you might find the max() function useful here.


#1.5 Produce a scatterplot of Canada's life expectancy vs. GDP per capita using ggplot2, without defining a new variable. That is, after filtering the gapminder data set, pipe it directly into the ggplot function. In your plot, put GDP per capita on a log scale. 


























































