---
title: "Consequences of Covid-19 for Candle Critiques"
author: "Tyler Allan"
date: "3/22/2021"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: lumen
---

For this assignment, you will produce a report exploring the impact of Covid-19 on scented candle ratings.

Write this assignment like a report.
You should describe the dataset, show appropriate plots, and interpret them.
Remove the instruction text from this document.

In your report, fold the code blocks, enable code highlighting, and set an HTML theme of your choice.
Set other document options as you prefer and as appropriate.


## The data

You will work with two datasets `Scented_all.xlsx` and `Unscented_all.xlsx`.
Each dataset contains the following information for the Top 3 selling scented or unscented candles sold on Amazon:
  
  1. `CandleID`: A numeric id variable for candle 1-3 in each category.
  2. `Date`: The date of the review.
  3. `Rating`: The 1-5 rating for the review.
  4. `Review`: For the scented candles, the text of the review.

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(gridExtra)
library(Hmisc)

Scented_all <- read_excel(here::here("Hw_05/Scented_all.xlsx"))
Unscented_all <- read_excel(here::here("Hw_05/Unscented_all.xlsx"))

```


## Part 1: Average daily ratings for scented and unscented candles

For each category of candle (scented versus unscented), for all days from **2017-01-01** onward, create a plot showing:

  1. Date on the x axis
    1. Hint: use `scale_x_date()`
    2. Format the date labels as `"%m-%Y"`
    3. Set the date breaks to be "6 months"
  2. **Average daily rating** of each candle on the y axis
  3. A smooth trend line (not linear; `method = "loess"`)
  4. A dashed vertical line at January 01, 2020
    1. Hint: use `geom_vline()` and set `xintercept = as.numeric(as.Date("2020-01-20"))`
    
Be sure to make the plots beautiful.

  1. Address overplotting of points using transparency or shapes.
  2. Use an appealing theme and color scheme.
  3. Set appropriately formatted axis titles and main plot title.
  4. Place the two plots side by side.
  
Note: The raw data are **individual reviews** so you will need to compute summary results to get average daily ratings for the desired date range.

Interpret these graphs.
What do they show about scented candles over the time period?

```{r}
Scented_Plot <- Scented_all %>%
  filter(Date > "2017-01-01") %>%
  group_by(Date) %>% 
  mutate(Avg_Rating = mean(Rating)) %>% 
  ggplot(
    aes(x = as.Date(Date),
        y = Avg_Rating
  )) + 
  geom_point(alpha = .5, color = "dark green", shape = 1) +
  scale_x_date(name = "Date", date_breaks = "6 months", date_labels = "%m-%Y") +
  geom_smooth(method = "loess", color = "gold") +
  geom_vline(xintercept = as.numeric(as.Date("2020-01-20")), linetype = "dashed") +
  scale_y_continuous(name = "Average Rating") +
  theme_minimal() +
  ggtitle("Average Scented Candle Reviews") + 
  guides(x = guide_axis(n.dodge = 2))

Unscented_Plot <- Unscented_all %>%
  filter(Date > "2017-01-01") %>%
  group_by(Date) %>% 
  mutate(Avg_Rating = mean(Rating)) %>% 
  ggplot(
    aes(x = as.Date(Date),
        y = Avg_Rating
  )) + 
  geom_point(alpha = .5, color = "dark green", shape = 1) +
  scale_x_date(name = "Date", date_breaks = "6 months", date_labels = "%m-%Y") +
  geom_smooth(method = "loess", color = "gold") +
  geom_vline(xintercept = as.numeric(as.Date("2020-01-20")), linetype = "dashed") + 
  scale_y_continuous(name = "Average Rating") + 
  theme_minimal() +
  ggtitle("Average Unscented Candle Reviews") +
  guides(x = guide_axis(n.dodge = 2))

grid.arrange(Scented_Plot, Unscented_Plot, ncol = 2)
```

### Description of the Plots

Both plots show a general skew towards the higher end of the scale meaning people are more likely to leave 5 star reviews than 1 star reviews. This trend begins to separate after the cutoff line place on January 20th 2020. Though the trend line for unscented candles does begin to trend down after January 20th, it does not appear to deviate enormously from what would be expected results. The average ratings for scented candles tell a different story. The trend line experiences a large drop off after January 20th 2020. This large decrease in average ratings could be related to individuals not realizing they have lost their sense of smell due to contracting COVID-19. 

## Part 2: Combined plot

Combine the two datasets into one and make a plot similar to above, but showing both unscented and scented candles.
Distinguish scented versus unscented candles using color, shape, and line type (for the trend lines).

```{r combine datasets,}
candles_full <- bind_rows(
	scented = Scented_all, 
	unscented = Unscented_all, 
	.id = "Scented"
)
```

```{r}
Combined_Plot <- candles_full %>%
  filter(Date > "2017-01-01") %>%
  group_by(Date, Scented) %>% 
  mutate(Avg_Rating = mean(Rating)) %>% 
  ggplot(
    aes(x = as.Date(Date),
        y = Avg_Rating,
        shape = Scented,
        color = Scented,
        fill = Scented
  )) + 
  geom_point(alpha = .3) +
  scale_x_date(name = "Date", date_breaks = "6 months", date_labels = "%m-%Y") +
  geom_smooth(method = "loess") +
  geom_vline(xintercept = as.numeric(as.Date("2020-01-20")), linetype = "dashed") +
  scale_y_continuous(name = "Average Rating") +
  theme_minimal() +
  ggtitle("Average Candle Reviews For Scented and Unscented Candles") + 
  guides(x = guide_axis(n.dodge = 2)) + 
  scale_color_viridis_d()

Combined_Plot
```

Which of these two sets of plots do you find easier to interpret?

The combined plot is easier to interpret specifically if you are trying to compare the reviews of unscented and scented candles. Having the data overlap and specifically having the trend lines on the same graph allows for a very clear comparison of the two average ratings.

## Part 3: Proportion of reviews mentioning "no scent"

Use the scented candles database and make one more plot showing:

  1. For **each month in 2019 and 2020**, 
  2. On the x axis, month and year
  3. On the y axis, the **proportion of reviews** mentioning that the scented candles have no smell
    - Include both the mean (proportion) and a 95% confidence interval
  4. Color the symbols by year (2019 vs 2020)

As before, make the 

Hints:

  1. To round each date to its months, use:
    - `mutate(month = lubridate::floor_date(Date, "month"))`
  2. To compute a mean and confidence interval for a variable on a plot, similar to a box plot, use:
    - `stat_summary(aes(y = noscent), fun.data = mean_cl_normal)`
  3. Use the function below to detect "no smell"-related text in the `Review` column.

```{r no scent function, include=FALSE}
str_no_scent <- function(x){
  as.numeric(
  	case_when(
  		str_detect(x, "[Nn]o scent") ~ "1", 
  		str_detect(x, "[Nn]o smell") ~ "1",
  		str_detect(x, "[Dd]oes not smell like") ~ "1",
  		str_detect(x, "[Dd]oesn't smell like") ~ "1",
  		str_detect(x, "[Cc]an't smell") ~ "1",
  		str_detect(x, "[Cc]annot smell") ~ "1",
  		str_detect(x, "[Ff]aint smell") ~ "1",
  		str_detect(x, "[Ff]aint scent") ~ "1",
  		str_detect(x, "[Dd]on't smell") ~ "1",
  		str_detect(x, "[Ll]ike nothing") ~ "1",
  		TRUE ~ "0"
  	)
  )
}

Scented_all$CodeReview <- str_no_scent(Scented_all$Review)
Scented_all$CodedReview_f <- factor(Scented_all$CodeReview, labels = c("scent", "noscent"))
print(Scented_all)
```

```{r}
Scented2_Plot <- Scented_all %>%
  filter(Date >= "2019-01-01") %>%
  mutate(month = lubridate::floor_date(Date, "month"),
         ) %>% 
  group_by(month , CodedReview_f) %>% 
  ggplot() +
  aes(x = month, y = (..count..)/sum(..count..), fill = CodedReview_f, color = month) +
  geom_bar(position = "dodge") +
  stat_summary(aes(y = CodedReview_f), fun.data = mean_cl_normal)+
  scale_color_viridis_d()

Scented2_Plot
```

What pattern do you see?
Is there a trend over time regarding smell-related reviews? 
How does 2020 compare to 2019?

There are more reviews that indicate that candles have a scent across the two years than reviews that suggest the candles have no scent. The proportion of reviews mentioning there was no scent to the candles increased from 2019 into 2020. while the total percentage is still lower than the percentage of reviews mentioning the scent of the candles the difference from 2019 to 2020 is noticable.  
