---
title: "Putting Plots Together"
editor_options: 
  chunk_output_type: console
---

## patchwork

```{r}
library(ggplot2)
library(patchwork)

p1 <- ggplot(mtcars) + 
  geom_point(aes(mpg, disp)) + 
  ggtitle('Plot 1')

p2 <- ggplot(mtcars) + 
  geom_boxplot(aes(gear, disp, group = gear)) + 
  ggtitle('Plot 2')

p3 <- ggplot(mtcars) + 
  geom_point(aes(hp, wt, colour = mpg)) + 
  ggtitle('Plot 3')

p4 <- ggplot(mtcars) + 
  geom_bar(aes(gear)) + 
  facet_wrap(~cyl) + 
  ggtitle('Plot 4')
```

patchwork is a very elegant package for composing multiple plots on a page. 
Let's check it out!

Use the `+` function to add multiple plots together:

```{r}
p1 + p2
```

If you save a patchwork as an object, you can create nested patchworks:

```{r}
patch <- p1 + p2
p3 + patch
```

You can control the layout of a patchwork using `()`, `|`, and `/`:

```{r}
p1 / p2

p1 | p2

p1 / (p2 | p3)
```

You can also control the grid with `plot_layout()`:

```{r}
p1 + p2 + p3 + plot_spacer() + p4 + plot_spacer()
  plot_layout(ncol = 3)
```

patchwork can collect guides for multiple plots:

```{r}
p1a <- ggplot(mtcars) + 
  geom_point(aes(mpg, disp, colour = mpg, size = wt)) + 
  ggtitle('Plot 1a')

p1a | (p2 / p3)

(p1a | (p2 / p3)) + plot_layout(guides = 'collect')
```

Finally, patchwork can add annotations and titles to the whole patchworkd:

```{r}
patchwork <- (p1 + p2) / p3
patchwork + plot_annotation(
  title = 'The surprising truth about mtcars',
  subtitle = 'These 3 plots will reveal yet-untold secrets about our beloved data-set',
  caption = 'Disclaimer: None of these plots are insightful'
)

patchwork + plot_annotation(tag_levels = 'A')
```

You can theme these annotations:

```{r}
patchwork + 
  plot_annotation(title = 'The surprising truth about mtcars',
                  subtitle = 'These 3 plots will reveal yet-untold secrets about our beloved data-set',
                  caption = 'Disclaimer: None of these plots are insightful',
                  theme = theme(
                    text = element_text(family = 'mono'),
                    plot.title = element_text(size = 18)
                  )
  )
```

patchwork can do a whole lot more too! Check out <https://patchwork.data-imaginist.com/articles/guides/assembly.html> and <https://patchwork.data-imaginist.com/articles/guides/layout.html>.
