---
title: "Gapminder annotation activity"
editor_options: 
  chunk_output_type: console
---

In this activity, imagine that you are a data analyst working for the prime minister of New Zealand.
The PM wants to know how trends in life expectancy in Zealand and its nearest neighbor, Australia, compare to other countries.

To answer this question, use the `gapminder` dataset.

1. Prepare your dataset. Load the `gapminder` package and make a new factor variable indicating whether a country is New Zealand/Australia (the `continent == "Oceania"` countries).

```{r}

```

2. Make a plot showing trendlines for life expectancy each country over time.
   Assign this plot to an object, then show it. (We will be adding more stuff to it later.)

```{r}

```

3. Now, update your plot above to emphasize the Oceania countries compared to the rest (e.g., use color, size [line width], and alpha).

```{r}

```

4. Label Australia and New Zealand on the plot.
   (Hint: You might need to update the data for this layer to filter out other countries and extra years.)
   
```{r}

```

5. Finish up your plot. Clean up the titles, axes, legend (do you need a legend?), theme, etc. Add any additional annotations you think would be helful to readers.

```{r}

```

6. As a supplement to the above plot, make a new **set** of plots comparing the Oceania countries to each of the other continents individually. Arrange these plots in a grid using `patchwork`.

```{r}

```

