# Homework 03

title: "HW03"
author: "Tyler Allan"
date: "2/18/2021"
output: 
    html_document: default 
    github_document: default 
    keep_md: TRUE


You will analyze data looking at the relationship between **green reputation** and three personality traits--**compassion**, **intellectual curiosity**, and **openness to experiences**.
The dataset includes data from **students** and **non-students**.

```{r}
dictionary <- readr::read_csv("green_dictionary.csv")
green_data <- readr::read_csv("green_data.csv")
library(tidyverse)
```

For your assignment, do the following.

  1. Inspect the item responses (e.g., with graphs or by summarizing distinct values). 
     Is anything unusual?

```{r}
print(green_data)
view(green_data)

summary(green_data$comp1)

green_data %>% 
  na_if("-99")


```

  2. Compute total scores for the four scales. 
     Recode variables as needed.
     
```{r}

```

  3. Rescale the variables so that they go from 0-100 instead of the original range.
     Name the recaled variables `*_pomp`.

```{r}

```

  4. Make plots that illustrate the distributions of the 4 POMP-scored variables.

```{r}

```

  5. Make scatterplots showing the relationships between **green reputation** and each personality trait.
     Include trend lines for **students** and **non-students**.
     What do these plots show?

```{r}

```

  6. Compare **green reputation** for students and non-students using a **rainfall plot** (bar + density + data points).

```{r}

```

  7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.
  
```{r}

```

  
In your assignment, prepare an RMarkdown file that includes both the coding and (for 4-7) narrative descriptions of your results.
