# Homework 03

title: "HW03"
author: "Tyler Allan"
date: "2/18/2021"
output: 
    html_document: default 
    github_document: default 
    keep_md: TRUE


You will analyze data looking at the relationship between **green reputation** and three personality traits--**compassion**, **intellectual curiosity**, and **openness to experiences**.
The dataset includes data from **students** and **non-students**.

```{r}
dictionary <- readr::read_csv("green_dictionary.csv")
green_data <- readr::read_csv("green_data.csv")
library(tidyverse)
```

For your assignment, do the following.

  1. Inspect the item responses (e.g., with graphs or by summarizing distinct values). 
     Is anything unusual?

```{r}
print(green_data)
summary(green_data$comp1)

new_green_data <- green_data %>% 
                na_if("-99")
print(new_green_data)
summary(new_green_data$green1)
```

  2. Compute total scores for the four scales. 
     Recode variables as needed.
     
```{r}
reversed <- c("green2", "comp6", "comp7", "comp8", "comp9", "comp10", 
              "intel7", "intel8", "intel9", "intel10", "open7", "open8", 
              "open9", "open10")
new_green_datar <- new_green_data %>% 
                    mutate(
                           across(all_of(reversed),
                           ~ recode(.x, "1" = 5, "2" = 4, "4" = 2, "5" = 1),
                           .names = "{.col}r")
                          ) 

new_green_datar %>% 
  rowwise() %>% 
  mutate(green_tot = sum(c(green1,green2r,green3,green4,green5), na.rm = TRUE),
         comp_tot = sum(c(comp1,comp2,comp3,comp4,comp5,comp6r,comp7r,comp8r,comp9r,comp10r), na.rm = TRUE),
         intel_tot = sum(c(intel1,intel2,intel3,intel4,intel5,intel6,intel7r,intel8r,intel9r,intel10r), na.rm = TRUE),
         open_tot = sum(c(open1,open2,open3,open4,open5,open6,open7r,open8r,open9r,open10r), na.rm = TRUE),) %>% 
  select(green_tot,comp_tot,intel_tot,open_tot)
```

  3. Rescale the variables so that they go from 0-100 instead of the original range.
     Name the rescaled variables `*_pomp`.

```{r}
new_green_datar <- new_green_datar %>% 
                    mutate(
                           across(all_of(reversed),
                           ~ recode(.x, "1" = 0, "2" = 25, "3" = 50, "4" = 75, "5" = 100),
                           .names = "{.col}_pomp")
                          ) 

```

  4. Make plots that illustrate the distributions of the 4 POMP-scored variables.

```{r}

```

  5. Make scatterplots showing the relationships between **green reputation** and each personality trait.
     Include trend lines for **students** and **non-students**.
     What do these plots show?

```{r}

```

  6. Compare **green reputation** for students and non-students using a **rainfall plot** (bar + density + data points).

```{r}

```

  7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.
  
```{r}

```

  
In your assignment, prepare an RMarkdown file that includes both the coding and (for 4-7) narrative descriptions of your results.
