# Homework 03

title: "HW03"
author: "Tyler Allan"
date: "2/18/2021"
output: 
    html_document: default 
    github_document: default 
    keep_md: TRUE


You will analyze data looking at the relationship between **green reputation** and three personality traits--**compassion**, **intellectual curiosity**, and **openness to experiences**.
The dataset includes data from **students** and **non-students**.

```{r}
dictionary <- readr::read_csv("green_dictionary.csv")
green_data <- readr::read_csv("green_data.csv")
library(tidyverse)
```

For your assignment, do the following.

  1. Inspect the item responses (e.g., with graphs or by summarizing distinct values). 
     Is anything unusual?

```{r}
print(green_data)
summary(green_data$comp1)

new_green_data <- green_data %>% 
                na_if("-99")
print(new_green_data)
summary(new_green_data$green1)
```

  2. Compute total scores for the four scales. 
     Recode variables as needed.
     
```{r}
reversed <- c("green2", "comp6", "comp7", "comp8", "comp9", "comp10", 
              "intel7", "intel8", "intel9", "intel10", "open7", "open8", 
              "open9", "open10")
new_green_datar <- new_green_data %>% 
                    mutate(
                           across(all_of(reversed),
                           ~ recode(.x, "1" = 5, "2" = 4, "4" = 2, "5" = 1),
                           .names = "{.col}r")
                          ) 

new_green_datar <- new_green_datar %>% 
  rowwise() %>% 
  mutate(green_tot = sum(c(green1,green2r,green3,green4,green5), na.rm = TRUE),
         comp_tot = sum(c(comp1,comp2,comp3,comp4,comp5,comp6r,comp7r,comp8r,comp9r,comp10r), na.rm = TRUE),
         intel_tot = sum(c(intel1,intel2,intel3,intel4,intel5,intel6,intel7r,intel8r,intel9r,intel10r), na.rm = TRUE),
         open_tot = sum(c(open1,open2,open3,open4,open5,open6,open7r,open8r,open9r,open10r), na.rm = TRUE),) 
```

  3. Rescale the variables so that they go from 0-100 instead of the original range.
     Name the rescaled variables `*_pomp`.

```{r}

new_green_data_pomp <- new_green_datar %>% 
                    mutate(
                           across(c(green1:open10r),
                           ~ recode(.x, "1" = 0, "2" = 25, "3" = 50, "4" = 75, "5" = 100),
                           .names = "{.col}_pomp")
                          )
new_green_data_pomp <- new_green_data_pomp %>% 
  rowwise() %>% 
  mutate(green_tot_pomp = sum(c(green1_pomp,green2r_pomp,green3_pomp,green4_pomp,green5_pomp), na.rm = TRUE),
         comp_tot_pomp = sum(c(comp1_pomp,comp2_pomp,comp3_pomp,comp4_pomp,comp5_pomp,comp6r_pomp,comp7r_pomp,
                               comp8r_pomp,comp9r_pomp,comp10r_pomp), na.rm = TRUE),
         intel_tot_pomp = sum(c(intel1_pomp,intel2_pomp,intel3_pomp,intel4_pomp,intel5_pomp,intel6_pomp,intel7r_pomp,
                                intel8r_pomp,intel9r_pomp,intel10r_pomp), na.rm = TRUE),
         open_tot_pomp = sum(c(open1_pomp,open2_pomp,open3_pomp,open4_pomp,open5_pomp,open6_pomp,open7r_pomp,
                               open8r_pomp,open9r_pomp,open10r_pomp), na.rm = TRUE),) 


```

  4. Make plots that illustrate the distributions of the 4 POMP-scored variables.

```{r}


ggplot(new_green_data_pomp[which(new_green_data_pomp$green_tot_pomp>0),]) +
  aes(x = green_tot_pomp) +
  geom_histogram(binwidth = 10)

ggplot(new_green_data_pomp[which(new_green_data_pomp$comp_tot_pomp>0),]) +
  aes(x = comp_tot_pomp) +
  geom_histogram(binwidth = 10)

ggplot(new_green_data_pomp[which(new_green_data_pomp$intel_tot_pomp>0),]) +
  aes(x = intel_tot_pomp) +
  geom_histogram(binwidth = 10)

ggplot(new_green_data_pomp[which(new_green_data_pomp$open_tot_pomp>0),]) +
  aes(x = open_tot_pomp) +
  geom_histogram(binwidth = 10)
```

#Response
Of the four pomp scored histograms only the green variable appears to be normally distributed around the center of the score distribution. The other three variables, open, intel, and comp appear to be skewed to the left such that more people responded in the higher range across these items.

  5. Make scatterplots showing the relationships between **green reputation** and each personality trait.
     Include trend lines for **students** and **non-students**.
     What do these plots show?

```{r}
new_green_data_pomp <- new_green_data_pomp %>%
  mutate(
    student = recode(student, "1" = "Nonstudent", "2" = "Student")
  )

ggplot(new_green_data_pomp[which(new_green_data_pomp$comp_tot_pomp>0),]) +
  aes(x = green_tot_pomp,
      y = comp_tot_pomp,
      fill = student,
      color = student) +
  geom_point() +
  geom_point() +
  geom_smooth(data = subset(new_green_data_pomp, student == "Nonstudent"| student == "Student"))

ggplot(new_green_data_pomp[which(new_green_data_pomp$intel_tot_pomp>0),]) +
  aes(x = green_tot_pomp,
      y = intel_tot_pomp,
      fill = student,
      color = student) +
  geom_point() +
  geom_point() +
  geom_smooth(data = subset(new_green_data_pomp, student == "Nonstudent"| student == "Student"))

ggplot(new_green_data_pomp[which(new_green_data_pomp$open_tot_pomp>0),]) +
  aes(x = green_tot_pomp,
      y = open_tot_pomp,
      fill = student,
      color = student) +
  geom_point() +
  geom_point() +
  geom_smooth(data = subset(new_green_data_pomp, student == "Nonstudent"| student == "Student"))

```
#Description
The trend lines for green reputation across all three personality variables appear to follow similar trends. Students low in green reputation tend to score below the nonstudents on their personality variables, but as green reputation scores increase, students tend to be higher in their persoanlity variables than nonstudents. 

  6. Compare **green reputation** for students and non-students using a **rainfall plot** (bar + density + data points).

```{r}

```

  7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.
  
```{r}

```

  
In your assignment, prepare an RMarkdown file that includes both the coding and (for 4-7) narrative descriptions of your results.
