---
title: "Beer Advocate: ABV and quality ratings"
author: "Your Name Here"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

For this activity, download the data file at this location to the `data` subfolder in your `participation` folder. 
This dataset contains information about different catagories of beer as rated on <beeradvocate.com>.
The data are originally from [here](http://scottjanish.com/relationship-of-abv-to-beer-scores/).

```{r}
url <- "https://raw.githubusercontent.com/USF-Psych-DataSci/Classroom/master/data/beer-advocate.csv"
```

The dataset contains 3 variables:

  1. Beer category
  2. Average alcohol by volume (ABV) in percentage (%)
  3. Average quality rating (rAVG) on a 1–5 scale
  
You should write a script to do the following:

  1. Download the dataset with an appropriate filename.

```{r}
if ( !dir.exists( here::here("participation", "data") ) ) {
  dir.create( here::here("participation", "data"), recursive = TRUE )
}

url <- "https://raw.githubusercontent.com/USF-Psych-DataSci/Classroom/master/data/beer-advocate.csv"
download.file(
  url, 
  here::here("participation", "data", "beeradv.csv"), 
  mode = "wb")
```

  2. Import the dataset into R.
     Be sure to check for any importing problems and correct as needed.
     
```{r}
beeradv <- read_csv( here::here("participation", "data", "beeradv.csv"),
                       na = c("", "NA", "Non-alcoholic")
                     )
beeradv




```  


  3. Make a figure illustrating the relationship between ABV and beer quality.
  
```{r}
ggplot(beeradv) +
    aes(x = `Average ABV`,
      y = rAVG) +
  geom_point()

```

  4. Compute a summary table with the mean and SD of each variable, along with their correlation.

```{r}
summarytable = beeradv %>% 
  summarize(across(c(`Average ABV`:rAVG),
            list(
              mu = ~mean(.x, na.rm = TRUE),
              sd = ~sd(.x, na.rm = TRUE)),
              r = cor(`Average ABV`, rAVG, use = "pairwise")
            ))
summarytable
```

  5. Describe the results.
